---
title: 'Module 11: Integer Programming'
output:
  pdf_document: default
  html_document: default
---

# Integer Programming Model

This report presents a small integer programming model and solves it using R.

We consider three decision variables:

- \(x_1\): number of units of Product 1
- \(x_2\): number of units of Product 2
- \(x_3\): number of units of Product 3

Each product generates a profit contribution per unit:

- Product 1: 40
- Product 2: 30
- Product 3: 50

The objective is to **maximize total profit** subject to a single resource constraint.

## Model Formulation

**Decision variables**

\[
x_1, x_2, x_3 \ge 0,\quad x_1, x_2, x_3 \in \mathbb{Z}
\]

**Objective function**

\[
\max Z = 40x_1 + 30x_2 + 50x_3
\]

**Constraint**

\[
2x_1 + x_2 + 3x_3 \le 5
\]

## R Code

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
# Install lpSolve if needed:
# install.packages("lpSolve")

library(lpSolve)
```

### Setting up the model in R

```{r model-setup}
# Coefficients for the objective function
objective <- c(40, 30, 50)   # x1, x2, x3

# Constraint matrix (each row is a constraint)
A <- matrix(c(2, 1, 3),      # 2*x1 + 1*x2 + 3*x3
            nrow = 1, byrow = TRUE)

# Direction of the constraints
dir <- c("<=")

# Right-hand side
rhs <- c(5)

# All three decision variables are integers
int.vec <- 1:3
```

### Solving the integer program

```{r solve-model}
solution <- lp(direction = "max",
               objective.in = objective,
               const.mat    = A,
               const.dir    = dir,
               const.rhs    = rhs,
               all.int      = TRUE)

solution$status          # 0 indicates optimal
solution$objval          # optimal objective value
solution$solution        # optimal values of x1, x2, x3
```

## Results

Based on the R solution, we obtain:

- Optimal objective value (maximum profit): **130**
- Optimal decision variable values:
  - \(x_1^\* = 1\)
  - \(x_2^\* = 3\)
  - \(x_3^\* = 0\)

These values satisfy the constraint

\[
2(1) + 1(3) + 3(0) = 5 \le 5
\]

and yield the maximum profit

\[
Z^\* = 40(1) + 30(3) + 50(0) = 130.
\]

## Summary

This document defined an integer programming model, implemented it in R using the `lpSolve` package, and reported the optimal solution and objective value.
